---
title: "Project 1"
author: "Libby Brill, Emi Degembe, Alvaro Ramos"
date: today
format: 
  html:
    code-fold: true
    code-tools: true
    toc: true
    embed-resources: true
    message: false
    warning: false
    theme: "darkly"
---

```{r}
#| echo: false
#| output: false

# quality of life
library(tidyverse)

# tables 
library(gt)

# json files
library(tidyjson)
library(jsonlite)

# webscraping/api
library(httr)

# modeling
library(nlme)

# visualizations
library(rnaturalearthdata)
library(rnaturalearth)
library(tidyverse)
library(ggtext)
library(scales)
library(fmsb)
```

```{r}
#| label: read in data from GitHub
#| output: false

cases_month <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-24/cases_month.csv')
cases_year <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-24/cases_year.csv')
```

```{r}
#| label: creating function with iteration for summary statistics 

########## HELPER FUNCTIONS ##########
######################################

# using the same summ_stats as practice activity 2
summ_stats <- function(x) {
    
    # if x contains numeric values
      # return a tibble with the mean, med, sd, IQR, and n observed
    if (is.numeric(x)) {
        out_df <- tibble(
            mean = mean(x, na.rm = TRUE), 
            median = median(x, na.rm = TRUE), 
            stdev = sd(x, na.rm = TRUE), 
            iqr = IQR(x, na.rm = TRUE),
            n = sum(!is.na(x)) # length won't work bc of NAs
        )
    # if x contains categorical, return tibble with # of levels in variable 
    } else if (is.factor(x) || is.character(x)) {
        out_df <- tibble(
            n_levels = n_distinct(x)
        )
    } else {
        stop("Inputted values are not numeric or categorical.")
    }

    #return the data frame.
    out_df
}

summarize_stats <- function(df, col = measles_lab_confirmed, mode = region) {
  ######### PARAMETERS ##########
  # input: df -- the dataframe of interest 
  # col: column of interest to summarize by: 
  #   measles_suspect, 
  #   measles_clinical, 
  #   measles_epi_linked, 
  #   measles_lab_confirmed [DEFAULT], 
  #   measles_total
  # mode: aggregate by options: 
  #   region [DEFAULT], 
  #   country, 
  #   iso3, 
  #   year 
  ###############################
  
  df |>
    select({{mode}}, {{col}}) |>
    group_by({{mode}}) |>
    #group_map removed the group column, so to retain that, group_modify.
    # See second example: https://dplyr.tidyverse.org/reference/group_map.html
    group_modify( ~ summ_stats(.x |>
                              pull({{col}}))) |>
    bind_rows()
}
```

```{r}
# API call
data <- GET("https://www.imf.org/external/datamapper/api/v1/NGDPDPC")

# creating function to make tibbles for each country from json data
country_values <- function(country){
  
  tibble(json = list(parse_json(data))) |> 
    unnest_wider(json) |> 
    unnest_longer(values) |> 
    unnest_wider(values) |>
    slice_head() |> 
    select({{country}}) |> 
    unnest_longer({{country}}) |> 
    rename(year = str_glue("{country}_id")) |> 
    pivot_longer(cols = {{country}},
                 names_to = "iso3",
                 values_to = "GDP_per_capita"
                 ) |> 
    mutate(year = as.numeric(year))
  
}

# storing countries available in both the cases_year dataset and json data
ids <- cases_year$iso3 |>
  as.factor() |>
  levels()

ids <- ids[ids %in% (
  tibble(json = list(parse_json(data))) |> 
    unnest_wider(json) |> 
    unnest_longer(values) |> 
    unnest_wider(values) |>
    slice_head() |> 
    names()
  )]

# creating dataset from json data for each country and merging to one
data <- map(.x = ids,
            .f = country_values
            ) |> 
  bind_rows()

# join to cases_year dataset to include all data
dataset <- cases_year |> 
  full_join(data) |> 
  filter(year >= 2012) #|> 
  #mutate(region_info = fct_recode(
    #region,
    #"Africa" = "AFRO",
    #"Americas" = "AMRO",
    #"Eastern Mediterranean" = "EMRO",
    #"Europe" = "EURO",
    #"South East Asia" = "SEARO",
    #"Western Pacific" = "WPRO"
    #)
    #)
```

```{r}
#| label: API call
data <- GET("https://ghoapi.azureedge.net/api/MCV2")

data <- tibble(json = list(parse_json(data))) |> 
  unnest_wider(json) |> 
  unnest_wider(value, names_sep = "_") |>
  select(-"@odata.context")
```

```{r}
#| label: creating function to extract data from json data
row <- function(v){
  data |> 
    select(v) |> 
    unnest_wider(v) |> 
    select(ParentLocationCode,
           #ParentLocation,
           SpatialDim,
           TimeDim,
           #Value,
           NumericValue
           )
}
```

```{r}
data <- map(.x = data |> names(),  
            .f = row
            ) |> 
  bind_rows() |> 
  rename(
    "region" = ParentLocationCode,
    #"region_info" = ParentLocation,
    "iso3" = SpatialDim,
    "year" = TimeDim,
    #value = Value,
    "mcv2" = NumericValue
  )
```

## Joining both sources to create one metadataset

```{r}
#| label: creating big metadata

metadata <- dataset |> 
  
  # join to cases_year dataset
  full_join(data |> select(-region),
            by = join_by(iso3, year)
            ) |> 
  filter(year >= 2012) |> 
  mutate(mcv2 = mcv2 / 100) |> 
  
  # recoding region to be more informative
  mutate(
    region_s = region,
    region = fct_recode(
      region,
      "Africa" = "AFRO",
      "Americas" = "AMRO",
      "Eastern Mediterranean" = "EMRO",
      "Europe" = "EURO",
      "South East Asia" = "SEARO",
      "Western Pacific" = "WPRO"
      )
    ) |> 
  
  mutate(p_lab_total = measles_lab_confirmed / measles_total)

colnames(metadata)

```

# Data Summaries & Visualizations

### Summary of Existing Variables

```{r}
summarize_stats(metadata |> filter(!(is.na(region))),
                measles_incidence_rate_per_1000000_total_population,
                region
                ) |> 
  ungroup() |> 
  arrange(-mean) |> 
  gt() |> 
  fmt_number(columns = -c(region, n), 
             decimals = 2) |> 
  cols_label("region" = "Region",
             "mean" = "Mean",
             "median" = "Median",
             "stdev" = "Standard Deviation",
             "iqr" = "IQR",
             "n" = "Sample Size",
             ) |> 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
    ) |> 
  tab_style(
    style = cell_text(style = "italic", decorate = "underline"),
    locations = cells_row_groups()
    ) |> 
  tab_header(
    title = md("Summary Statistics for Total Number of Measles Cases for Each
               1 Million Individuals")
  ) |> 
  tab_caption(
    caption = md("**Table X** Summary table visualizing various statistics of
                 the number of total measles cases for each 1 million individuals at the region level.
                 Samples are collected annually from 2012 to 2024 for each
                 region. Notably, some region demonstrate large differences
                 between the mean and median.")
    ) |> 
  data_color(
    columns = "mean",
    method = "numeric",
    palette = c("#CCEEDF", "#BAE8D4", "#A8E2CA",
                "#96DCBF", "#83D6B5", "#6ED0AB",
                "#57CAA1", "#3BC397", "#00BD8D"
                )
    )
```

Using the table to see and compare the number of total measles cases for each 1 million individuals between the different regions, we see that the region averaging the largest number seems to be the Western Pacific and the region averaging the smallest number seems to be the Americas. Addressing the research question, Western Pacific overall seems to demonstrate the largest number of total measles cases. Looking further into each region, another table is created to see the number of total measles cases for each 1 million individuals for the top 5 countries of each region.

```{r}
metadata |> 
  filter(!(is.na(region)),
         !(is.na(country))
         ) |> 
  
  # used to assign each country to the appropriate region when merging
  select(region, country) |> 
  unique() |>
  arrange(country) |> 
  
  full_join(summarize_stats(metadata |> filter(!(is.na(region)), !(is.na(country))),
                            measles_incidence_rate_per_1000000_total_population,
                            country)) |> 
  group_by(region) |> 
  slice_max(mean, n = 5) |> 
  arrange(-mean) |> 
  gt() |> 
  fmt_number(columns = -c(country, region, n), 
             decimals = 2) |> 
  cols_label("country" = "Country",
             "mean" = "Mean",
             "median" = "Median",
             "stdev" = "Standard Deviation",
             "iqr" = "IQR",
             "n" = "Sample Size",
             ) |> 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
    ) |> 
  tab_style(
    style = cell_text(style = "italic", decorate = "underline"),
    locations = cells_row_groups()
    ) |> 
  tab_header(
    title = md("Summary Statistics for Total Number of Measles Cases for Each
               1 Million Individuals")
  ) |> 
  tab_caption(
    caption = md("**Table X** Summary table visualizing various statistics of
                 the number of total measles cases for each 1 million individuals.
                 Shown are the top 5 countries for each region.
                 Samples are collected annually from 2012 to 2024 for each
                 country. Notably, some countries demonstrate large differences
                 between the mean and median.")
    ) |> 
  data_color(
    columns = "mean",
    method = "numeric",
    palette = c("#CCEEDF", "#BAE8D4", "#A8E2CA",
                "#96DCBF", "#83D6B5", "#6ED0AB",
                "#57CAA1", "#3BC397", "#00BD8D"
                )
    )
```

Using this table to look at how the number of total measles cases for each 1 million individuals differs for the different top 5 countries for each region, we see that for the Western Pacific region, Mongolia demonstrates a mean number that seems to be substantially larger than any of the countries for any region.

# Statistical Model

Our second research question of interest was "are there associations between measles and region resources such as GDP?" Below, we fit a generalized least squares (GLS) model using region, GDP per capita, and vaccination percentages as explanatory variables to address the research question. Since the model uses yearly data, the model utilizes an autocorrelation structure for the errors using year as the covariate.

```{r}
m <- gls(
  measles_incidence_rate_per_1000000_total_population ~ region + log_GDP_per_capita + mcv2,
  data = metadata |> mutate(log_GDP_per_capita = log(GDP_per_capita)),
  na.action = na.omit,
  correlation = corAR1(form = ~ year | region / country)
  )

summary(m)$tTable |> 
  as_tibble(rownames = "var") |> 
  mutate(region = c("Quantitative",
                    "Region (Africa as baseline) (overall p-value = 0.057)",
                    "Region (Africa as baseline) (overall p-value = 0.057)",
                    "Region (Africa as baseline) (overall p-value = 0.057)",
                    "Region (Africa as baseline) (overall p-value = 0.057)",
                    "Region (Africa as baseline) (overall p-value = 0.057)",
                    "Quantitative",
                    "Quantitative"
                    )
         ) |> 
  mutate(var = var |> 
           str_remove("region") |> 
           str_replace("log_GDP_per_capita", "log(GDP per capita ($))") |> 
           str_replace("mcv2", "MCV2")
         
         ) |> 
  group_by(region) |> 
  gt() |> 
  fmt_number(columns = c("Value", "Std.Error", "t-value"), 
              decimals = 3
             ) |> 
  fmt_number(columns = c("p-value"), 
              decimals = 3
             ) |> 
  cols_align(align = "left", columns = region) |>
  cols_label("var" = "Variable",
             "Value" = "Estimated coefficient"
             ) |> 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
    ) |> 
  tab_style(
    style = cell_text(style = "italic", decorate = "underline"),
    locations = cells_row_groups()
    ) |> 
  tab_header(
    title = md("Estimated coefficients for modeled variables "
               )
    ) |> 
  tab_caption(
    caption = md("**Table X** Estimated explanatory variable coefficients for a model using number of measles cases per million population as the dependent variable. Model utilizes an autocorrelation structure for the errors using year as the covariate."
                 )
    )
```

Based on the output of the model, the only statistically significant variable seems to be log(GDP per capita) as the associated p-value is 0.003. Estimated coefficient of the variable is −28.2, which may broadly be interpreted as richer countries seeing less measles cases per million. Regarding the second research question, the model suggests that there does seem to be an association between measles and region resources such as GDP.